---
#title: "Grade overview"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(xlsx)
library(reshape)
library(ggplot2)
library(plyr)
library(dplyr)
library(sqldf)
library(knitr)
library(xtable)



ThisYear<-"17/18" #school year
numOfRowsToImport<-83
numOfRowsEvalToImport<-65

semester<-2
OddSemester<-TRUE
show_semester2 <-FALSE; show_semester1 <-FALSE; 
ifelse(semester==1,show_semester1<-TRUE,show_semester2<-TRUE)
# show_semester2 <- TRUE
# show_semester1 <- FALSE
OddSems<-seq(1,9,by=2)
EvenSems<-seq(2,10,by=2)

file <- loadWorkbook("/Users/hendrik/OneDrive - Aalborg Universitet/02 semester coordination/grades exam Over The years Mea Med1 and 2 Medialogy.xlsx")
fileName<-"/Users/hendrik/OneDrive - Aalborg Universitet/02 semester coordination/grades exam Over The years Mea Med1 and 2 Medialogy.xlsx"
#df1 <- readColumns(getSheets(file)[[1]], startColumn = 1, endColumn = 24, startRow = 1, endRow = numOfRowsToImport, header = T)
df<-read.xlsx(fileName, sheetIndex=1, sheetName=NULL, rowIndex=NULL,
              startRow=1, endRow=numOfRowsToImport, colIndex=seq(1:17),
              as.data.frame=TRUE, header=TRUE, colClasses=NA,
              keepFormulas=FALSE, encoding="unknown",check.names=FALSE)

#read in the course evaluations
dfe<-read.xlsx(fileName, sheetIndex=3, sheetName=NULL, rowIndex=NULL,
              startRow=1, endRow=numOfRowsEvalToImport, colIndex=seq(1:11),
              as.data.frame=TRUE, header=TRUE, colClasses=NA,
              keepFormulas=FALSE, encoding="unknown",check.names=FALSE)

dfm<-melt(df,id=c("location","education","year","sem","course","exam"))
dfem<-melt(dfe,id=c("location","education","year","sem","course","item"))
dfem$variable<-as.character(dfem$variable)
dfem[dfem$variable=="very good",]$variable<-"1"
dfem[dfem$variable=="good",]$variable<-"2"
dfem[dfem$variable=="medium",]$variable<-"3"
dfem[dfem$variable=="not so good",]$variable<-"4"
dfem[dfem$variable=="not good",]$variable<-"5"
dfem$variable<-as.numeric(dfem$variable)
#identify courses that are pass/fail in a given year / careful later on if switches occur in study plan
coursesWithPassFail<-sqldf("select distinct year, education, course from dfm where variable='I' and value>=-1")
coursesWithPassFail$isPassFail<-TRUE
dfm<-merge(dfm,coursesWithPassFail,all.x = TRUE)

#remove pass/fails from grades courses (grades courses) and grades from pass/fails
#dfmNPF<-dfm[(is.na(dfm$isPassFail) & !(dfm$variable %in% c('B','I'))),]
#dfmPF<-dfm[!(is.na(dfm$isPassFail)) & dfm$variable %in% c('U','EB','B','I'),]
dfm<-dfm[((is.na(dfm$isPassFail) & !(dfm$variable %in% c('B','I')))) | (!(is.na(dfm$isPassFail)) & dfm$variable %in% c('U','EB','B','I')),]

#remove grades from pass/fail courses
#dfm<-dfm[is.na(dfm$isPassFail) | dfm$variable %in% c('U','EB','B','I'),]

#dfm<-dfm[!is.na(dfm$value),]

#replace empty entries with 0 and letters (EB,U,I,B) with numbers 
dfm[is.na(dfm$value),]$value<-0
ifelse(nrow(is.na(dfem$value)) >0,dfem[is.na(dfem$value),]$value<-0)

dfm$variable<-as.character(dfm$variable)
dfm[dfm$variable=="EB",]$variable<-"-5"
dfm[dfm$variable=="U",]$variable<-"-4"
dfm[dfm$variable=="I",]$variable<-"0"
dfm[dfm$variable=="B",]$variable<-"1"


dfm<-dfm[dfm$variable!="pass" & dfm$variable!="fail",]
#remove all grades from passfail courses that are not pass/fail grades
# dfm[dfm$variable=="pass",]$variable<-"2"
# dfm[dfm$variable=="fail",]$variable<-"0"

dfm$variable<-as.numeric(dfm$variable)


df[df$sem==semester & df$exam=="exam" & df$course=="PI",]

columns = names(dfm)[seq(1:6)]

#colu = names(data)[-3]

# dfx <- dfm %>%
#   group_by_at(vars(one_of(columns))) %>%
#   mutate(percent=value/sum(value))

dfmp<-sqldf("select location, education, year, sem, course, exam,  sum(value) as total from dfm group by  location, education, year, sem, course, exam")
dfemp<-sqldf("select location, education, year, sem, course, item,  sum(value) as total from dfem group by  location, education, year, sem, course, item")

dfm<-merge(dfm,dfmp,all.x = TRUE)
dfem<-merge(dfem,dfemp,all.x = TRUE)

dfm$percent<-round(dfm$value/dfm$total,digits=2)
dfem$percent<-round(dfem$value/dfem$total,digits = 2)
MaxGradesPercentYear<-ceiling(max(dfm[dfm$year==ThisYear & dfm$exam=="exam" & dfm$sem==semester,]$percent)*10)/10+0.1
MaxQualPercentYear<-ceiling(max(dfem[dfem$year==ThisYear & dfem$exam=="quality" & dfem$sem==semester,]$percent)*10)/10+0.1

dfm<-dfm[!is.na(dfm$variable),]


```

## Grade distribution comparison across modules ThisYear
```{r fig4, echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}
ggplot(dfm[dfm$sem==semester & dfm$exam=="exam" & !(dfm$variable %in% c("pass", "fail")) ,],aes(x=variable,y=percent,color=course,shape=course))+
  theme_bw()+ ggtitle(paste("Grade distribution comparison on Med2",ThisYear)) +
  geom_line(data=dfm[dfm$sem==semester & dfm$exam=="exam" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+ 
  scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfm[dfm$sem==semester & dfm$exam=="exam" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+
  #geom_text(data=dfm[dfm$sem==semester & dfm$exam=="exam" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+
scale_x_continuous(labels=c("EB","U","-3","0/I","B","2","4","7","10","12"), breaks = c(-5,-4,-3,0,1,2,4,7,10,12))+xlab("grades")

```

```{r fig4a, echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}
ggplot(dfem[dfem$sem==semester & dfm$item=="quality" ,],aes(x=variable,y=percent,color=course,shape=course))+
  theme_bw()+ ggtitle(paste("Course evaluation on Med2",ThisYear)) +
  geom_line(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$year  ==ThisYear ,])+ 
  scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$year  ==ThisYear ,])+ scale_x_reverse(labels=c("bad","not so good","average","good","very good"), breaks = c(5,4,3,2,1))+xlab("quality of course")

```



```{r fig8, echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester1}
ggplot(dfm[dfm$sem==semester & dfm$exam=="exam" ,],aes(x=variable,y=percent,color=course,shape=course))+
  theme_bw()+ ggtitle(paste("Grade distribution comparison on Med1 ",ThisYear)) +
  geom_line(data=dfm[dfm$sem==semester & dfm$exam=="exam"  & dfm$year  ==ThisYear ,])+ 
  scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$year  ==ThisYear ,])+
  #geom_text(data=dfm[dfm$sem==semester & dfm$exam=="exam" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+
scale_x_continuous(labels=c("EB","U","-3","0/I","B","2","4","7","10","12"), breaks = c(-5,-4,-3,0,1,2,4,7,10,12))+xlab("grades")
```


<!-- courses<-ifelse(semester==1,c("AVS","GPRO/IP","PBL","P1"),c("PI","MMA","ID","P2")) -->
<!-- coursesLong<-ifelse(semester==1,c("Audio-Visual Sketching","Introduction to Programming (GPRO)","PBL","P1 project""),c("Programming for Interaction","Mathematics for Multimedia Applications","Interaction Design","P2 project")) -->
## Grade distributions of modules and comparison with recent years
Med2 module grade distributions and evaluations in comparison to previous years (grey smoothed error bands indicating .95 confidence intervals of the percentage of students attaining that grade or providing that rating). 
Absolute numbers of attained grades or ratings are printed in black in the figures below.


```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}
# courses<-ifelse(semester==1,c("AVS","GPRO/IP","PBL","P1"),c("PI","MMA","ID","P2"))
# coursesLong<-ifelse(semester==1,c("Audio-Visual Sketching","Introduction to Programming (GPRO)","PBL","P1 project"),c("Programming for Interaction","Mathematics for Multimedia Applications","Interaction Design","P2 project"))
# 
ggplot(dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="PI" & !(dfm$variable %in% c("pass", "fail")) ,],aes(x=variable,y=percent))+
  theme_bw()+ ggtitle("Programming for Interaction") +
  geom_line(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="PI" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+ 
  scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="PI" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+
  geom_text(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="PI" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+ 
scale_x_continuous(labels=c("EB","U","-3","0/I","B","2","4","7","10","12"), breaks = c(-5,-4,-3,0,1,2,4,7,10,12))+xlab("grades")
```

```{r table1, echo=FALSE,message=F, warning=F, results='asis',eval=show_semester2}
library(xtable)
options(xtable.comment = FALSE)
print(xtable(df[df$sem==semester & df$course=="PI" & df$exam=="exam" & df$year==ThisYear,c(7:8,11:17)], type="latex", row.names = FALSE,digits=c(0,0,0,0,0,0,0,0,0),include.rownames=FALSE,comment=FALSE))

#knitr::opts_chunk$set(echo = TRUE, fig.align="right")
#![](MED2A.tiff) 

```

```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}
ggplot(dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="PI",],aes(x=variable,y=percent))+
  theme_bw()+ 
  geom_line(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="PI" & dfem$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+   scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="PI"& dfem$year  ==ThisYear ,])+
  geom_text(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="PI" & dfem$year ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+ 
scale_x_reverse(labels=c("bad","not so good","average","good","very good"), breaks = c(5,4,3,2,1))+xlab("quality of course")+ylab("percentage of responses")

```

<!-- ```{r table1, echo=FALSE,message=F, warning=F, results='asis',eval=show_semester2} -->



<!-- #knitr::opts_chunk$set(echo = TRUE, fig.align="right") -->
<!-- #![](MED2A.tiff) -->

<!-- ``` -->

<!-- ```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2} -->
<!-- #nrow(dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="PI" & !(dfm$variable %in% c("pass", "fail")) ,]) -->

<!-- ``` -->


```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}
ggplot(dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="MMA" & !(dfm$variable %in% c("pass", "fail")) ,],aes(x=variable,y=percent))+ theme_bw()+ ggtitle("Mathematics for Multimedia Applications (MMA)") +   geom_line(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="MMA" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+scale_x_continuous(labels=c("EB","U","-3","0/I","B","2","4","7","10","12"), breaks = c(-5,-4,-3,0,1,2,4,7,10,12))+xlab("grades") + scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="MMA" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+
  geom_text(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="MMA" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+xlab("grades")
```

```{r table2, echo=FALSE,message=F, warning=F, results='asis',eval=show_semester2}
library(xtable)
print(xtable(df[df$sem==semester & df$course=="MMA" & df$exam=="exam" & df$year  ==ThisYear,c(7:8,11:17)], type="latex", row.names = FALSE,digits=c(0,0,0,0,0,0,0,0,0,0)),include.rownames=FALSE,comment=FALSE)

#knitr::opts_chunk$set(echo = TRUE, fig.align="right")
#![](MED2A.tiff) 

```

```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}

ggplot(dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="MMA",],aes(x=variable,y=percent))+
  theme_bw()+ #ggtitle("Programming for Interaction (PI)") +
  geom_line(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="MMA" & dfem$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+   scale_y_continuous(labels = scales::percent,breaks = seq(0,MaxGradesPercentYear,0.1))+
  geom_point(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="MMA" & dfem$year  ==ThisYear ,])+
  geom_text(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="MMA" & dfem$year ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+ 
scale_x_reverse(labels=c("bad","not so good","average","good","very good"), breaks = c(5,4,3,2,1))+xlab("quality of course")+ylab("percentage of responses")
```


```{r echo=FALSE,message=F,warning=FALSE, out.width = '80%',eval=show_semester2}
ggplot(dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="ID" & !(dfm$variable %in% c("pass", "fail")) ,],aes(x=variable,y=percent))+
  theme_bw()+ ggtitle("Interaction Design (ID)") +
  geom_line(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="ID" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+scale_x_continuous(labels=c("EB","U","-3","0/I","B","2","4","7","10","12"), breaks = c(-5,-4,-3,0,1,2,4,7,10,12))+xlab("grades")+ 
  scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="ID" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+
  geom_text(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="ID" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+xlab("grades")
```

```{r table3, echo=FALSE,message=F, warning=F, results='asis',eval=show_semester2}
library(xtable)
print(xtable(df[df$sem==semester & df$course=="ID" & df$exam=="exam" & df$year  ==ThisYear,c(7:8,11:17)], type="latex", row.names = FALSE,digits=c(0,0,0,0,0,0,0,0,0,0)),include.rownames=FALSE,comment=FALSE)

#knitr::opts_chunk$set(echo = TRUE, fig.align="right")
#![](MED2A.tiff) 

```

```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}
ggplot(dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="ID",],aes(x=variable,y=percent))+
  theme_bw()+ #ggtitle("Programming for Interaction (PI)") +
  geom_line(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="ID" & dfem$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+   scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="ID" & dfem$year  ==ThisYear ,])+
  geom_text(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="ID" & dfem$year ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+ 
scale_x_reverse(labels=c("bad","not so good","average","good","very good"), breaks = c(5,4,3,2,1))+xlab("quality of course")+ylab("percentage of responses")
```


```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}
ggplot(dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="P2" & !(dfm$variable %in% c("pass", "fail")) ,],aes(x=variable,y=percent))+
  theme_bw()+ ggtitle("P2 Project") +
  geom_line(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="P2" &  !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+scale_x_continuous(labels=c("EB","U","-3","0/I","B","2","4","7","10","12"), breaks = c(-5,-4,-3,0,1,2,4,7,10,12))+xlab("grades") + scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="P2" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,])+
  geom_text(data=dfm[dfm$sem==semester & dfm$exam=="exam" & dfm$course=="P2" & !(dfm$variable %in% c("pass", "fail")) & dfm$year  ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+xlab("grades")
```

```{r table4, echo=FALSE,message=F, warning=F, results='asis',eval=show_semester2}
library(xtable)
print(xtable(df[df$sem==semester & df$course=="P2" & df$exam=="exam" & df$year  ==ThisYear,c(7:8,11:17)], type="latex", row.names = FALSE,digits=c(0,0,0,0,0,0,0,0,0,0)),include.rownames=FALSE,comment=FALSE)

#knitr::opts_chunk$set(echo = TRUE, fig.align="right")
#![](MED2A.tiff) 

```
```{r echo=FALSE,message=F, warning=F, out.width = '80%',eval=show_semester2}

ggplot(dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="P2",],aes(x=variable,y=percent))+
  theme_bw()+
  geom_line(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="P2" & dfem$year  ==ThisYear ,])+ 
  geom_smooth(se=TRUE)+   scale_y_continuous(labels = scales::percent)+
  geom_point(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="P2" & dfem$year  ==ThisYear ,])+
  geom_text(data=dfem[dfem$sem==semester & dfem$item=="quality" & dfem$course=="P2" & dfem$year ==ThisYear ,],aes(label=value),hjust=.5, vjust=1.5)+ 
scale_x_reverse(labels=c("bad","not so good","average","good","very good"), breaks = c(5,4,3,2,1))+xlab("quality of course")+ylab("percentage of responses")
```

